---
title: "PRojet MRR"
author: "Pierre"
date: '2022-11-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(readxl)
library(MASS)
library(corrplot)
library(kernlab)
library(glmnet)
library(FactoMineR)
library(dplyr)
library(Amelia)
library(ggplot2)
library(plotly)
library(lars)
```

```{r, include = FALSE}
Tab <- read_excel('Vect.xlsx')
vec_depth <- read_excel('depth.xlsx')
head(Tab)
```

```{r}
plot(Tab$`'CHL 4'`,Tab$`'THERM 4'`, xlab = "Concentration CHL1 in [ng/l]", ylab = "Température")
```

```{r}
par(mfrow= c(2,2))
hist(Tab$`'THERM 3'`, breaks = 20, main = "Histogram of THERM 3", xlab = "Degrees")
hist(Tab$`'THERM 7'`, breaks = 20, main = "Histogram of THERM 7", xlab = "Degrees")
hist(Tab$`'THERM 10'`, breaks = 20, main = "Histogram of THERM 10", xlab = "Degrees")
hist(Tab$`'THERM 13'`, breaks = 20, main = "Histogram of THERM 13", xlab = "Degrees")
```


```{r}
par(mfrow= c(2,2))
hist(Tab$`'CHL 5'`, main = "Histogram of CHL5", breaks= 20, xlab = "CHL concentrations in [ng/l]")
hist(Tab$`'CHL 8'`, main = "Histogram of CHL8", breaks= 20, xlab = "CHL concentrations in [ng/l]")
hist(Tab$`'CHL 10'`, main = "Histogram of CHL10", breaks= 20, xlab = "CHL concentrations in [ng/l]")
hist(Tab$`'CHL 13'`, main = "Histogram of CHL13", breaks= 20, xlab = "CHL concentrations in [ng/l]")

```
As we can see, we may need to transformate our data. We can see for CHL5 and CHL8 that the values are mostly into the first part of the histogram. \
We can modelise those values by putting a log fonction into it.

```{r}
Tab_temp <- apply(Tab[,c(23:35)], 2, log)
Tab_temp <- data.frame(Tab_temp)

par(mfrow= c(2,2))
hist(Tab_temp$X.CHL.5., main = "Histogram of CHL5", breaks= 20, xlab = "CHL concentrations in [ng/l]")
hist(Tab_temp$X.CHL.7., main = "Histogram of CHL8", breaks= 20, xlab = "CHL concentrations in [ng/l]")
hist(Tab_temp$X.CHL.10., main = "Histogram of CHL10", breaks= 20, xlab = "CHL concentrations in [ng/l]")
hist(Tab_temp$X.CHL.13., main = "Histogram of CHL13", breaks= 20, xlab = "CHL concentrations in [ng/l]")

```
After we use the log fonction on the CHL data we saw that the distribution is now more even. 

# BATS

%    i      | no. of 5 day step
%    k      | no. year
%    CHLORO | DCHL + NCHL  in [ng/l].
%    SSH    | sea surface elevation
%    CC     | cloud cover
%    WS     | wind speed
%    SR     | shortwave radiation
%    Therm  | vertical thermic profile

Discontinuité -> 5 days la seule a modifier

```{r}
bats_y = 33.75912    
bats_x = -64.02191

Cos_Days <- function(x){
  cos(2*pi*(x/73));
}

Sin_Days <- function(x){
  sin(2*pi*(x/73));
}

BATS <- Tab[Tab$`'latitude'` >= 33 & Tab$`'latitude'` <= 33.8 & Tab$`'longitude'` >= -64.5 & Tab$`'longitude'` <= -64,]

BATS <- BATS[order(BATS$`'year'`),]

BATS_2005 <- BATS[BATS$`'year'` == 2005,]
BATS_2005 <- as.data.frame(scale(BATS_2005))

days_5_cos <- Cos_Days(BATS_2005$`'5days'`)
days_5_sin <- Sin_Days(BATS_2005$`'5days'`)

BATS_2005$days_5_cos = c(days_5_cos)
BATS_2005$days_5_sin = c(days_5_sin)

lm5days <- lm(BATS_2005$`'CHL 5'`~ BATS_2005$`'THERM 1'` + BATS_2005$`'CHL 1'` + BATS_2005$`'SSH'` + BATS_2005$`'CC'` + BATS_2005$`'WS'` + BATS_2005$`'SR'` + BATS_2005$`'5days'`+BATS_2005$days_5_cos+BATS_2005$days_5_sin, data = BATS_2005)
summary(lm5days)
plot(lm5days)
attributes(res)
coefficients(res)


QD_err_mean <- function(m, o){
  sqrt(mean((m - o)^2))
}

RMSE_log <- function(m,o){
  sqrt(mean(10^(m)- 10^(log(o)))^2)
}

RMSE <- function(m, o){
  sqrt(mean((m - o)^2))
}

B <- predict.lm(lm5days, newdata = as.data.frame(BATS_2005$`'CHL 5'`))

RMSE_log_rsult <- RMSE_log(B, BATS_2005$`'CHL 5'`)

QDEM_coeff_BATS_2005 <- QD_err_mean(B, BATS_2005$`'CHL 5'`)

QDEM_mean_BATS_2005 <- mean(BATS_2005$`'CHL 5'`)

QDEM_stat_BATS_2005 <- QD_err_mean(B,BATS_2005$`'CHL 5'`)*100/mean(BATS_2005$`'CHL 5'`)

shapiro.test(lm5days$residuals)


#moche donc faut affiner -> enlever des variables ajoute pour enlever les variables corrélées entre elle

```
R2 = 0.83 so 83% of the variables are explained by this dataset with a 17% mistakes.

#Correlation Matrix 
We add the 2 new columns to the dataframe.
We only save the data that we have (SSH, WS, SR, TH1, CHL1, CC, year, 5days, Lat, Long)
```{r}
First_tab <- data.frame(   Tab$`'SSH'`, Tab$`'CC'`, Tab$`'WS'`, Tab$`'SR'`, Tab$`'THERM 1'`, Tab$`'CHL 1'`, Tab$`'5days'`,  Tab$`'latitude'`, Tab$`'longitude'`)



corTab <- cor(First_tab)
corrplot(corTab, method="color")
```


#Missing Values
```{r}

BATS_2005 <- BATS[BATS$`'year'` == 2005,]
BATS_2005 <- as.data.frame(scale(BATS_2005))

#Compter le nombre de valeurs manquantes dans chacune des colonnes
sapply(BATS_2005,function(x) sum(is.na(x)))

#Compter le nombre de NA dans une colonne en particulier
sum(is.na(BATS_2005$columns))

#Afficher les lignes dont la valeur dans la colonne "colonne" est NA
BATS_2005[is.na(BATS_2005$columns),]


missmap(BATS_2005, main = "Missing values vs observed")
```

# Fonction print 
```{r}
vect = Tab
depth = vec_depth


BATS <- Tab[Tab$`'latitude'` >= 33 & Tab$`'latitude'` <= 33.8 & Tab$`'longitude'` >= -64.5 & Tab$`'longitude'` <= -64,]
BATS <- BATS[order(BATS$`'year'`, BATS$`'5days'`),]

g_bis <- function( df){
   sub_df = as.data.frame(df)
   d = c()
   for (i in sub_df$`'5days'`){
      if (i >= 1 && i <= 6){
         d[i] = "Jan"
      }
      if (i >= 7 && i <= 12){
         d[i] = "Feb"
      }
      if (i >= 13 && i <= 18){
         d[i] = "Mar"
      }
      if (i >= 19 && i <= 24){
         d[i] = "Apr"
      }
      if (i >= 25 && i <= 30){
         d[i] = "May"
      }
      if (i >= 31 && i <= 36){
         d[i] = "Jun"
      }
      if (i >= 37 && i <= 42){
         d[i] = "Jul"
      }
      if (i >= 43 && i <= 48){
         d[i] = "Aug"
      }
      if (i >= 49 && i <= 54){
         d[i] = "Sep"
      }
      if (i >= 55 && i <= 60){
         d[i] = "Oct"
      }
      if (i >= 61 && i <= 66){
         d[i] = "Nov"
      }
      if (i >= 67 && i <= 73){
         d[i] = "Dec"
      }
   }
   sub_df
   tmp4 = sub_df[1:73, 10]
   plot_ly(y = depth, x = d, z = t(tmp4), type='contour', colors = "Spectral", reversescale = TRUE, x = c(0, -50)) %>% layout(title = "The Chlorophyll-a values of the NEMO-PISCES for different years", yaxis = list(title = "Depth (m)"), showlegend= FALSE)
}


g <- function(year, BATS){
   sub_df = subset(BATS, year==year)
   d = c()
   for (i in sub_df$`'5days'`){
      if (i >= 1 && i <= 6){
         d[i] = "Jan"
      }
      if (i >= 7 && i <= 12){
         d[i] = "Feb"
      }
      if (i >= 13 && i <= 18){
         d[i] = "Mar"
      }
      if (i >= 19 && i <= 24){
         d[i] = "Apr"
      }
      if (i >= 25 && i <= 30){
         d[i] = "May"
      }
      if (i >= 31 && i <= 36){
         d[i] = "Jun"
      }
      if (i >= 37 && i <= 42){
         d[i] = "Jul"
      }
      if (i >= 43 && i <= 48){
         d[i] = "Aug"
      }
      if (i >= 49 && i <= 54){
         d[i] = "Sep"
      }
      if (i >= 55 && i <= 60){
         d[i] = "Oct"
      }
      if (i >= 61 && i <= 66){
         d[i] = "Nov"
      }
      if (i >= 67 && i <= 73){
         d[i] = "Dec"
      }
   }
   sub_df
   tmp4 = sub_df[1:73, 23:40]
   plot_ly(y = depth, x = d, z = t(tmp4), type='contour', colors = "Spectral", reversescale = TRUE, x = c(0, -50)) %>% layout(title = "The Chlorophyll-a values of the NEMO-PISCES for different years", yaxis = list(title = "Depth (m)"), showlegend= FALSE)
}
 g(2000, BATS)
 
 
```




#Régression logistique Forward.

```{r}

BATS_2005 <- BATS[BATS$`'year'` == 2005,]
BATS_2005 <- as.data.frame(scale(BATS_2005))

days_5_cos <- Cos_Days(BATS_2005$`'5days'`)
days_5_sin <- Sin_Days(BATS_2005$`'5days'`)

BATS_2005$days_5_cos = c(days_5_cos)
BATS_2005$days_5_sin = c(days_5_sin)

BATS_2005 <- BATS_2005[colnames(BATS_2005)[c(1, 2, 3, 4, 5, 23, 27, 41,45,46)]]
BATS_2005$`'CHL 5'` <- as.factor(BATS_2005$`'CHL 5'`)

res = glm(`'CHL 5'`~., data = BATS_2005, family = binomial)
summary(res)
attributes(res)
coefficients(res)

resall<-glm(`'CHL 5'`~.,data=BATS_2005,family="binomial");
res0<-glm(`'CHL 5'`~ 1,data=BATS_2005,family=binomial);

resfor<-step(res0,list(upper=resall),direction='forward')
print(resfor)
##Régression logistique Backward

resback<-step(res,direction='backward')
print(resback)

#Régression logistique Stepwise

resstep<-step(res,direction='both');
print(resstep)

formula(resstep)
formula(resback)
formula(resfor)

sum((resfor$residuals)^2)
sum((resback$residuals)^2)
sum((resstep$residuals)^2)
```


#Lasso Method with Normal Data
```{r}
BATS_2005 <- BATS[BATS$`'year'` == 2005,]
BATS_2005 <- as.data.frame(scale(BATS_2005))

days_5_cos <- Cos_Days(BATS_2005$`'5days'`)
days_5_sin <- Sin_Days(BATS_2005$`'5days'`)

BATS_2005$days_5_cos = c(days_5_cos)
BATS_2005$days_5_sin = c(days_5_sin)

BATS_2005 <- BATS_2005[colnames(BATS_2005)[c(1, 2, 3, 4, 5, 23, 27, 41,45,46)]]

dim <- dim(BATS_2005)

n <- dim[1]
p <- dim[2]

#BATS_2005$`'CHL 5'` <- as.factor(BATS_2005$`'CHL 5'`)

X=as.matrix(subset(BATS_2005, select=-`'CHL 5'`))
Y=BATS_2005$`'CHL 5'`

#k-fold cross validation
model <- cv.glmnet(X, Y, alpha = 1)
best_lambda <- model$lambda.min
plot(model)
 
best_model <- glmnet(X, Y, alpha = 1, lambda = best_lambda)
coef(best_model)

modlasso <- lars(X, as.numeric(Y),type="lasso")

fit_lasso <- as.matrix(predict.lars(modlasso, X, type="fit", mode="lambda", s=best_lambda)$fit)

RSS <- 0
for(i in 1:length(fit)){
  RSS = RSS + (fit[i] - Y[i])^2
}
Model_lasso <- cbind(X, fit_lasso)
Model_lasso <- as.data.frame(Model_lasso)
g_bis(Model_lasso)

library("lattice")

levelplot(depth,fit_lasso, Model_lasso)

coef(modlasso)
RSS

RMSE_Lasso <- RMSE(fit_lasso, BATS_2005$`'CHL 5'`)
par(mfrow=c(1,2))
plot(modlasso)
plot(c(modlasso$lambda, 0), pch=16, type="b", col="blue")

"AIC : "
-n * log(RSS / n) + 2 * p
```


#Ridge with Normal Dataset
```{r}
BATS_2005 <- BATS[BATS$`'year'` == 2005,]
BATS_2005 <- as.data.frame(scale(BATS_2005))

days_5_cos <- Cos_Days(BATS_2005$`'5days'`)
days_5_sin <- Sin_Days(BATS_2005$`'5days'`)

BATS_2005$days_5_cos = c(days_5_cos)
BATS_2005$days_5_sin = c(days_5_sin)

BATS_2005 <- BATS_2005[colnames(BATS_2005)[c(1, 2, 3, 4, 5, 23, 27, 41,45,46)]]

my_lambda = seq(0,10,0.01)

RIDGE_LAMBDA <- lm.ridge(BATS_2005$`'CHL 5'`~., data = BATS_2005, lambda = my_lambda )
plot(RIDGE_LAMBDA)

chosen_lambda <- which.min(RIDGE_LAMBDA$GCV)
my_lambda[chosen_lambda]

#k-fold cross validation
model_ridge <- cv.glmnet(X, Y, alpha = 0)
best_lambda <- model_ridge$lambda.min
plot(model_ridge)
 
best_model_ridge <- glmnet(X, Y, alpha = 0, lambda = best_lambda)
coef(best_model_ridge)

RIDGE_CHOSEN_LAMBDA  <- lm.ridge(BATS_2005$`'CHL 5'`~., data = BATS_2005, lambda = best_lambda)
coefridge <-  coef (lm.ridge(BATS_2005$`'CHL 5'`~., data = BATS_2005, lambda = best_lambda))

Yridge=as.matrix(BATS_2005)%*%as.vector(coef(RIDGE_CHOSEN_LAMBDA))
plot(Yridge)

RIDGE_FINAL <- lm.ridge(BATS_2005$`'CHL 5'`~., data = BATS_2005, lambda = best_lambda )

X <- as.matrix(BATS_2005[,-7])
Y <- as.matrix(BATS_2005[7])

modridge <- glmnet(X,Y,alpha = 0 , lambda = chosen_lambda)

fit_ridge <- as.matrix(predict(modridge, newdata=BATS_2005, s=0, type = "coefficients",interval = "confidence"))


RSS_ridge <- 0
for(i in 1:length(fit_ridge)){
  RSS_ridge = RSS_ridge + (fit_ridge[i] - Y[i])^2
}
dim(fit_ridge)
- n* log(RSS_ridge / n) + 2 * p

#RMSE_Ridge <- RMSE(fit_ridge, BATS_2005$`'CHL 5'`)
```


#Creation of Spatial + Temporal Dataset

```{r}
bats_y = 32.10851
bats_x = -64.01247

BATS <- Tab[Tab$`'latitude'` >= 33 & Tab$`'latitude'` <= 33.8 & Tab$`'longitude'` >= -64.5 & Tab$`'longitude'` <= -64,]

BATS <- BATS[order(BATS$`'year'`),]

days_5_cos <- Cos_Days(days5)
days_5_sin <- Sin_Days(days5)

Tab$days_5_cos = c(days_5_cos)
Tab$days_5_sin = c(days_5_sin)

BATS_8_points <- Tab[Tab$`'year'` == 1992 & Tab$`'5days'`== 1 ,]
Lat_8_points <- BATS_8_points$`'latitude'`
Long_8_point <- BATS_8_points$`'longitude'`


Lat_long_8_pts <- cbind(Lat_8_points, Long_8_point)
plot(Lat_long_8_pts, type = "p", main = "BATS and 8 others points")



coord_1 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[1,1],]
coord_2 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[2,1],]
coord_3 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[3,1],]
BATS <- Tab[Tab$`'latitude'` == Lat_long_8_pts[4,1],]
coord_5 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[5,1],]
coord_6 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[6,1],]
coord_7 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[7,1],]
coord_8 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[8,1],]
coord_9 <- Tab[Tab$`'latitude'` == Lat_long_8_pts[9,1],]

coord_1 <- coord_1[colnames(coord_1)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
coord_2 <- coord_2[colnames(coord_2)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
coord_3 <- coord_3[colnames(coord_3)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
BATS <- BATS[colnames(BATS)[c(1, 2, 3, 4, 5, 23, 27, 45, 46)]]
coord_5 <- coord_5[colnames(coord_5)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
coord_6 <- coord_6[colnames(coord_6)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
coord_7 <- coord_7[colnames(coord_7)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
coord_8 <- coord_8[colnames(coord_8)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
coord_9 <- coord_9[colnames(coord_9)[c(1, 2, 3, 4, 5, 23, 45, 46)]]


coord_1_CHL <- coord_1$`'CHL 1'`
coord_1_TH  <- coord_1$`'THERM 1'`


Final_Tab <- cbind(coord_1, coord_2, coord_3, BATS, coord_5, coord_6, coord_7, coord_8, coord_9)

for (i in 1:8){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".1")
}
for (i in 9:16){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".2")
}
for (i in 17:24){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".3")
}
for (i in 34:41){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".5")
}
for (i in 42:49){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".6")
}
for (i in 50:57){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".7")
}
for (i in 58:65){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".8")
}
for (i in 66:73){
  names(Final_Tab)[i]=paste(names(Final_Tab[i]),".9")
}
Final_Tab

```
#Creation of Temporal Dataset
```{r}
BATS_2 <- Tab[Tab$`'latitude'` >= 31 & Tab$`'latitude'` <= 33 & Tab$`'longitude'` >= -65 & Tab$`'longitude'` <= -63,]



BATS_2 <- BATS_2[order(BATS_2$`'year'`),]

BATS_2004 <- BATS_2[BATS_2$`'year'` == 2004,]
BATS_2005 <- BATS_2[BATS_2$`'year'` == 2005,]
BATS_2006 <- BATS_2[BATS_2$`'year'` == 2006,]


coord_2004<-BATS_2004[colnames(BATS_2004)[c(1, 2, 3, 4, 5, 23, 45, 46)]]
coord_2005<-BATS_2005[colnames(BATS_2005)[c(1, 2, 3, 4, 5, 23, 27, 45, 46)]]
coord_2006<-BATS_2006[colnames(BATS_2006)[c(1, 2, 3, 4, 5, 23, 45, 46)]]


for(i in 1:8){
    names(coord_2004)[i]=paste(names(coord_2004)[i],".1")
}
for(i in 1:8){
    names(coord_2006)[i]=paste(names(coord_2006)[i],".3")
}

Final_tab_temp<-cbind(coord_2004,coord_2005,coord_2006)

Final_tab_temp
```

Add the 8 others points to the dataframe. And then train/test/validation with the year
Choose 1992 -> 2000 -> 2004 -> 2006
```{r}
Tab_train <- Tab[Tab$`'year'` >= 1992 & Tab$`'year'` < 1998,]
Tab_test  <- Tab[Tab$`'year'` >= 1998 & Tab$`'year'` < 2003,]
Tab_valid  <- Tab[Tab$`'year'` >= 2003 & Tab$`'year'` <= 2008,]

Tab_train <- Tab_train[order(Tab_train$`'year'`) ,]
Tab_test <- Tab_test[order(Tab_test$`'year'`) , ]
Tab_valid <- Tab_valid[order(Tab_valid$`'year'`) , ]
```


#Elastic Net
```{r}
BATS_2005 <- BATS[BATS$`'year'` == 2005,]
BATS_2005 <- as.data.frame(scale(BATS_2005))

days_5_cos <- Cos_Days(BATS_2005$`'5days'`)
days_5_sin <- Sin_Days(BATS_2005$`'5days'`)

BATS_2005$days_5_cos = c(days_5_cos)
BATS_2005$days_5_sin = c(days_5_sin)

BATS_2005 <- BATS_2005[colnames(BATS_2005)[c(1, 2, 3, 4, 5, 23, 27, 41,45,46)]]

X=as.matrix(subset(BATS_2005, select=-`'CHL 5'`))
Y=BATS_2005$`'CHL 5'`

alpha <- c(0,0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)
par(mfrow = c(2,6))
for (i in alpha){
  elastic_net <- cv.glmnet(x = X,y = Y, alpha = i, nfolds = 10)
  elastic_net.min <- glmnet(X,Y, alpha = i, lambda = elastic_net$lambda.min)
  print(elastic_net.min$lambda)
  plot(elastic_net.min$beta)
  grid()
}
```


#PCA
```{r}
BATS <- Tab[Tab$`'latitude'` >= 33 & Tab$`'latitude'` <= 33.8 & Tab$`'longitude'` >= -64.5 & Tab$`'longitude'` <= -64,]

BATS <- BATS[order(BATS$`'year'`),]

BATS_2005 <- BATS[BATS$`'year'` == 2005,]
BATS_2005 <- as.data.frame(scale(BATS_2005))

days_5_cos <- Cos_Days(BATS_2005$`'5days'`)
days_5_sin <- Sin_Days(BATS_2005$`'5days'`)

BATS_2005$days_5_cos = c(days_5_cos)
BATS_2005$days_5_sin = c(days_5_sin)

BATS_2005 <- BATS_2005[colnames(BATS_2005)[c(1, 2, 3, 4, 5, 23, 27, 41,45,46)]]


BATS_2005_PCA <- PCA(BATS_2005, graph = F)
eigvalues <- data.frame(BATS_2005_PCA$eig)
plot(eigvalues$percentage.of.variance)
X_PCA <- BATS_2005_PCA$ind$coord



fviz_pca_biplot(BATS_2005_PCA)

fviz_contrib(BATS_2005_PCA, choice = "var",axes = 1:2,  top = 10 )

library(factoextra)
fviz_eig(BATS_2005_PCA, addlabels = TRUE)
plot(BATS_2005_PCA)
fviz_pca_ind(BATS_2005_PCA, pointsize = "cos2", pointshape = 21, fill = "#E7B800", repel = TRUE)
```



